Data Predictive Control with Random Forests:\\Theory and Applications\tnoteref{mytitlenote}

### 2017-09-29 paper/cs-dpc.tex
completed section 5 and fixed figures problems
completed abstract, section 5 and conclusion
Added flow scheme, motivation. Implemented comments from Italy.

The fundamental reason why DPC  is suitable for such a problem is that when the complexity rises, there is a huge cost to model all the states given by the dynamical system \eqref{E:bilinear1}. For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states. Thus, DPC is based solely on one state of the model i.e. the zone temperature that can be easily measured with a thermostat. This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively. Therefore, $\tY_{\mathrm{k+j|k}}:=x_{\mathrm{k+j|k}}^1$, where $x^1$ is the first component of $x$.
Next, we define the non-manipulated features $\tX^d_{\mathrm{k|k}}$. At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{R}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
$\tX^d_{\mathrm{k|k}}:=[ d_{\mathrm{k+j-N|k}} ,\dots,d_{\mathrm{k+j-1|k}}, x_{\mathrm{k|k}}^1,\dots, x_{\mathrm{k-\delta|k}}^1 ]$, where $\delta$ is the order of autoregression.
Finally, the inputs in DPC are exactly same as in MPC. i.e. $\tX^c_{\mathrm{k+j-1|k}}:=u_{\mathrm{k+j-1|k}}$.

@@

\subsubsection{Optimization} 
\label{SSS:dpc_opt}
For a fair comparison with MPC, we cast DPC optimization problem as follows:
\begin{subequations}
\begin{align}
\text{min } & \sum_{j=1}^{N} {\tY}_{\mathrm{k+j|k}} \mathcal{Q}^{(1,1)} {\tY}_{\mathrm{k+j|k}} + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \alpha_j^T \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T \label{SE:dpc}\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{align} \label{E:dpc}
\end{subequations}
\noindent Here $\alpha = \beta$ for DPC-RT and $\alpha = \hat{\Theta}$ for DPC-En.
Note that, \eqref{E:dpc} is DPC analog of \eqref{E:mpc}. The only difference is the state dynamics \eqref{SE:mpc1} and \eqref{SE:mpc2} are now replaced with \eqref{SE:dpc}.

@@

We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Fig.~\ref{F:validation}. It is visible how trees have a high variance, and the forests are more accurate. Note that data from January and May was not used for training. The quantitative summary of the accuracy is given in Tab.~\ref{T:validation}. We can see that the random forests are better in all respects.
\begin{table}[h!]
	\centering
	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
	\captionsetup{justification=centering}
	\begin{tabular}{c|c|c|c}
		\toprule
		& RMSE & $R^2$ score & EV  \\ 
		\midrule
		tree-$\tY_{\mathrm{k+1|k}}$    &  0.42 &  0.75 & 0.76    \\
		tree-$\tY_{\mathrm{k+6|k}}$  & 0.64 &  0.41  & 0.42 \\
		forest-$\tY_{\mathrm{k+1|k}}$  & 0.29 & 0.87  & 0.88 \\
		forest-$\tY_{\mathrm{k+6|k}}$  & 0.38 & 0.78 & 0.80 \\
		\bottomrule
	\end{tabular}
	\label{T:validation}
\end{table}

>>>>

The fundamental reason why DPC  is suitable for such a problem is that when the complexity rises, there is a huge cost to model all the states given by the dynamical system \eqref{E:bilinear1}. For example, states in the bilinear model also include slab temperatures which require modeling of structural and material properties in detail and often we also need to install new sensors to capture additional states. Thus, DPC is based solely on one state of the model, i.e. the zone temperature that can be easily measured with a thermostat. This serves as the output variable $\tY$ of interest for which we build $N$ trees and $N$ forests as described in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, respectively. Therefore, $\tY_{\mathrm{k+j|k}}:=x_{\mathrm{k+j|k}}^1$, where $x^1$ is the first component of $x$.
Next, we define the non-manipulated features $\tX^d_{\mathrm{k|k}}$. At time $k$, for the tree $\mathcal{T}_j$ and the forest $\mathcal{R}_j$, we base these features to include weather disturbances, external disturbances due to occupancy and equipments, and autoregressive terms of the room temperature, i.e.
$\tX^d_{\mathrm{k|k}}:=[ d_{\mathrm{k+j-N|k}} ,\dots,d_{\mathrm{k+j-1|k}}, x_{\mathrm{k|k}}^1,\dots, x_{\mathrm{k-\delta|k}}^1 ]$, where $\delta$ is the order of autoregression.
Finally, the inputs in DPC are exactly same as in MPC. i.e. $\tX^c_{\mathrm{k+j-1|k}}:=u_{\mathrm{k+j-1|k}}$.

@@

\subsubsection{Optimization} 
\label{SSS:dpc_opt}
For a fair comparison with MPC, we cast DPC optimization problem as follows:
\begin{subequations}
\begin{align}
\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) \mathcal{Q}^{(1,1)} ({\tY}_{\mathrm{k+j|k}} - x_{ref}) + c^T \tX^c_{\mathrm{k+j-1|k}}+  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \alpha_j^T \left[1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} \right]^T \label{SE:dpc}\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{align} \label{E:dpc}
\end{subequations}
\noindent Here $\alpha = \beta$ for DPC-RT and $\alpha = \hat{\Theta}$ for DPC-En.
Note that, \eqref{E:dpc} is DPC analog of \eqref{E:mpc}. The only difference is the state dynamics \eqref{SE:mpc1} and \eqref{SE:mpc2} are now replaced with \eqref{SE:dpc}.

@@

We compare the prediction for the first time step $\tY_{\mathrm{k+1|k}}$ and the 6-hour ahead prediction $\tY_{\mathrm{k+6|k}}$ for a week in the month of May in Figure~\ref{F:validation}. It is visible how trees have a high variance, and the forests are more accurate. Note that data from January and May were not used for training. The quantitative summary of the accuracy is given in Table~\ref{T:validation}. We can see that the random forests are better in all respects.
\begin{table}[h!]
	\centering
	\begin{tabular}{cccc}
		\toprule
		& RMSE & $R^2$ score & EV  \\ 
		\midrule
		tree-$\tY_{\mathrm{k+1|k}}$    &  0.42 &  0.75 & 0.76    \\
		tree-$\tY_{\mathrm{k+6|k}}$  & 0.64 &  0.41  & 0.42 \\
		forest-$\tY_{\mathrm{k+1|k}}$  & 0.29 & 0.87  & 0.88 \\
		forest-$\tY_{\mathrm{k+6|k}}$  & 0.38 & 0.78 & 0.80 \\
		\bottomrule
	\end{tabular}
	\caption{Quantitative comparison of root mean square error (RMSE), $R^2$ score, and explained variance (EV) for trees and forests for different predictions steps.}
	\captionsetup{justification=centering}
	\label{T:validation}
\end{table}

### 2017-09-29 paper/dpc.tex
Added flow scheme, motivation. Implemented comments from Italy.

Using separation of variables, the training process is divided into two steps. \textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity. It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion. \textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$. We have validated this linear model assumption in \cite{Jain2016}. As we shall see in Sec.~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program. This process is illustrated in Fig.~\ref{F:dpc-sepvars}.
\begin{figure}[t!]
	\centering
	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
	\caption{Separation of variables. \textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features. Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control. \textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
	\captionsetup{justification=centering}
	\label{F:dpc-sepvars}
\end{figure}

@@

While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients. We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest. Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known. Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
\begin{align}
\begin{aligned}
\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:dpcrf}
\end{align}
DPC-En is graphically described in Fig.~\ref{F:dpc-algo-rf}. The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control. In the next section, we compare DPC-RT and DPC-En to MPC for a building model.

>>>>

Using separation of variables, the training process is divided into two steps. \textit{Step 1:} The trees and the ensembles are trained only on $\tX^d$, which eases the computational complexity. It is important to note that besides external disturbances, $\tX^d$ also contains autoregressive terms of the output $\tY$ which is the main reason for the state space explosion. \textit{Step 2:} Linear regression models are trained in the leaves (or terminal nodes) of the trees which are function of only $\tX^c$. We have validated this linear model assumption in \cite{Jain2016}. As we shall see in Section~\ref{SS:dpcrt} and \ref{SS:dpcrf}, the second step reduces the run-time control problem into a convex program. This process is illustrated in Figure~\ref{F:dpc-sepvars}.
\begin{figure}[t!]
	\centering
	\includegraphics[width=20pc]{figures/dpc-sepvars.eps}
	\caption{Separation of variables. \textit{Step 1:} Tree $\mathcal{T}_1$ is trained only on the disturbances $\tX_d$ as the features. Tree $\mathcal{T}_2$ uses both the disturbances $\tX_d$ and the control variables $\tX_c$ for splitting and is thus not computationally suitable for control. \textit{Step 2:} In the leaf $R_i$ of the trees, a linear regression model parametrized by $\beta_i$ is defined as a function only of the control variables.}
	\captionsetup{justification=centering}
	\label{F:dpc-sepvars}
\end{figure}

@@

While the offline training burden in DPC-En is slightly increased compared to DPC-RT, in the control step we exploit the better accuracy, and lower variance properties of the random forest. 
If a forest has $t$ number of trees, given the forecast of disturbances, we have $t$ sets of linear coefficients. We simply average out all the coefficients from all the trees to get one linear model represented by $\hat{\Theta}_j$ for each forest. Note that the averaging step can only be done in run-time because the leaf of each tree can be narrowed down only when the $\tX^d$ is known. Thus, for $N$ forests, we again have exactly $N$ linear equality constraints in the optimization problem below:
\begin{align}
\begin{aligned}
\text{min } & \sum_{j=1}^{N} ({\tY}_{\mathrm{k+j|k}})^2 \mathcal{Q} + {\tX^c}^T_{\mathrm{k+j-1|k}} \mathcal{R} {\tX^c}_{\mathrm{k+j-1|k}} +  \lambda\epsilon_j\\
\text{s.~t. } & \ \ \ \ \ \tY_{\mathrm{k+j|k}} =  \hat{\Theta}^T_j [1,\tX^c_{\mathrm{k|k}},\dots,\tX^c_{\mathrm{k+j-1|k}} ]^T \\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tX}^c \leq \tX^c_{\mathrm{k+j-1|k}} \leq \bar{\tX}^c\\ 
& \ \ \ \ \ \ \ \ \ \ \ \ \underline{\tY}-\epsilon_j \leq \tY_{\mathrm{k+j|k}} \leq \bar{\tY} + \epsilon_j\\\
& \ \ \ \ \ \ \ \ \ \ \ \ \ \ \epsilon_j \geq 0, \ j = 1,\dots,N.
\end{aligned}
\label{E:dpcrf}
\end{align}
DPC-En is graphically described in Figure~\ref{F:dpc-algo-rf}. The ensemble data predictive control (DPC-En) is the first such method to bridge the gap between ensemble predictive models (such as random forests) and receding horizon control. In the next section, we compare DPC-RT and DPC-En to MPC for a building model.

### 2017-09-29 paper/intro.tex
awesome figure
Added flow scheme, motivation. Implemented comments from Italy.

Consider a multivariable dynamical system subject to external disturbances. The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system. With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance. In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances. The caveat is that MPC requires a reasonably accurate physical representation of the system. This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.

@@

%	\noindent (1) \textbf{Model complexity:} 
%	Building high fidelity models for intelligent control can be both cost and time prohibitive. For example, in the case of buildings, this requires installation of thousands of new sensors, expert domain knowledge and several months to years of efforts depending upon the size and complexity of structure of the building. We present this case-study in detail in Sec.~\ref{S:casestudy}.

The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.

In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

\subsection{Main contributions and paper organization}
In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control. In this paper, we extend these results providing the following contributions:
\begin{enumerate}
	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
	\begin{itemize}
		\item DPC with regression trees, and
		\item DPC with random forests.
	\end{itemize}
	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland. We show that DPC captures 70\% variance in MPC and offers a comparable performance.
	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house. We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
\end{enumerate}

%In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control. We now formally present the algorithms for DPC and apply them to a practical building model identified using real weather data and sensor measurements. We present a comprehensive case study to demonstrate the applicability of DPC to demand response, especially how DPC can provide financial incentives to the end-customers bypassing the need for high fidelity models. 
%In particular, we first describe two underlying algorithms: DPC with regression trees and DPC with random forests in Sec.~\ref{S:dpc}. 
%In Sec.~\ref{S:proof}, we apply DPC to a bilinear building model, whose parameters were identified using experiments on a building in Switzerland, to show a one-to-one comparison of receding horizon control with DPC against a benchmark MPC controller. In Sec.~\ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a multi-story 19 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.

@@

%We begin with description of DPC-RT and DPC-En algorithms, and Gaussian Processes for DPC in Sec.~\ref{S:dpc}. Sec.~\ref{S:casestudy} presents a case study on DR using a bilinear building model. Here, we also validate the black-box models used for control, and compare DPC against MPC. We discuss the challenges associated with DPC in Sec.~\ref{SS:challenges}. We conclude the paper with a summary of the results and a brief discussion on the future work in Sec.~\ref{S:conclusion}.

>>>>

\subsection{Motivation}
Consider a multivariable dynamical system subject to external disturbances. The first and foremost requirement for making any decision is to obtain the underlying control-oriented predictive model of the system. With a reasonable forecast of the external disturbances, these models should predict the state of the system in the future and thus Model Predictive Control (MPC) can act preemptively to provide a desired system behavior while optimizing a desired performance. In particular, MPC has been proven to be very powerful for multivariable systems in the presence of input and output constraints, and forecast of the disturbances. The caveat is that MPC requires a reasonably accurate physical representation of the system. This makes MPC unsuitable for control of complex plants such as natural gas processing, oil refineries, boilers, manufacturing plants, and buildings where the user expertise, time, and associated sensor costs required to develop a model are very high \cite{Sturzenegger2016,vzavcekova2014}.

@@

%	\noindent (1) \textbf{Model complexity:} 
%	Building high fidelity models for intelligent control can be both cost and time prohibitive. For example, in the case of buildings, this requires installation of thousands of new sensors, expert domain knowledge and several months to years of efforts depending upon the size and complexity of structure of the building. We present this case-study in detail in Section~\ref{S:casestudy}.

\paragraph{Example} As an example of such modeling complexity, consider the grey-box approach in \cite{Braun2002}. The scope of such approach is to predict the heat transfer rate to the air within the building. In particular, the authors addressed a single zone modeling problem using an equivalent $RC$ model. To this aim, five different types of structures are considered in the example: external walls, ceiling/roof, floor, internal walls and windows. Each of these elements are represented using $3$ resistances, $2$ capacitances and $2$ temperatures, except for the windows that are represented only with resistances since they have a negligible energy storage. A model consisting of $8$ states and $9$ inputs, including disturbances such as solar radiation and others, with the instantaneous heat gain to the building air from all surfaces as output, is created for the zone. To have a complete LTI model, $9$ parameters needed to be estimated through a non-trivial training process, using a collection of information associated with a physical description of the building (see \cite{Braun2002} for more details). In Section \ref{S:realCaseStudy} we will consider an application of the methodology proposed in this paper on a real house with $10$ different zones. This means that, in order to have a model of the zones as the one just described, only for what concerns the heat transfer rate within the building, and hence neglecting the energy consumption model of the house, we would need to consider a model with around $80$ states and $90$ parameters to be estimated.

\paragraph{Objective} The question now is, can we employ data-driven techniques to reduce the cost of modeling and still exploit the benefits that MPC has to offer? Can we build automatic and data-driven approaches for control purposes that are also adaptive, scalable and interpretable? In this paper we address this problem introducing \textit{Data Predictive Control (DPC)} to bridge the gap between Machine Learning and Predictive Control.

\subsection{Related work}
In the literature, several studies deal with the data-driven control techniques using machine learning algorithms \cite{Hou2013}. For example, Artificial Neural Networks are exploited to create data-driven models to be used as plant simulator in closed-loop with a Supervisory Model Predictive Control in \cite{Afram2017}. In \cite{Macarulla2017} the authors proposed a predictive control strategy based on Neural Networks, for boilers control in buildings, to decide the optimal time to switch-on the plan to guarantee energy saving and thermal comfort. However, the strategy is not easily scalable to different types of plants and does not use optimization theory in the closed-loop scheme. In \cite{Costanzo2016} a reinforcement learning control strategy, called Model-Assisted Batch Reinforcement Learning, is considered to provide data-driven control for the demand response problem in HVAC systems. Reinforcement Learning is a model-free methodology and for this reason it can not be used for on-line optimization in a closed-loop predictive control scheme. In \cite{Ferreira2012} the authors considered a data-driven predictive control based on Neural Networks to guarantee energy saving and  thermal comfort in public buildings. Neural Networks are used in the closed-loop control scheme to determine a thermal comfort index based on parameters that can be measured or estimated, but no  system dynamics with internal state are included into the optimization problem. More papers related to this topic can be found in the literature, but to the best of the authors' knowledge none of them address the problem of including data-driven state models in the optimal predictive control loop, and hence allow to set up a MPC-like optimal control problem, as we do in DPC.

\subsection{Main contributions and paper organization}
In our previous work \cite{Behl2016,Jain2016,JainACC2017,JainCDC2017}, we introduced the concept of DPC for receding horizon control. In this paper, we extend these results providing the following contributions.
\begin{enumerate}
	\item In Section \ref{S:dpc}, we formally present the two following control techniques:
	\begin{enumerate}
		\item DPC with regression trees, and
		\item DPC with random forests.
	\end{enumerate}
	\item In Section \ref{S:proof}, we demonstrate the strength of DPC for receding horizon control via one-to-one comparison against a benchmark MPC controller using a bilinear building model whose parameters were identified using experiments on a building in Switzerland. We show that DPC captures 70\% variance in MPC and offers a comparable performance.
	\item Section \ref{S:casestudy} describes a practical application of DPC for Demand Response, where we apply DPC to a 6 story 22 zone building model in EnergyPlus \cite{Crawley2001} for which model-based control is not economical and practical due to extreme complexity. We show scalability and efficiency of DPC in providing financial incentives to the end-customers bypassing the need for high fidelity models. We observe that DPC provides the desired power reduction with an average error of 3\%.
	\item In Section \ref{S:realCaseStudy}, we implement DPC on the real data from an off-grid house located in L'Aquila (Italy), to find the optimal ON/OFF scheduling for the heating system in order to save energy while guaranteeing thermal comfort for the occupants. We quantify the total amount of energy saved, with respect to the classical bang-bang controller widely used in houses for temperature control, using an EnergyPlus model built specifically for the house. We show we can perform an energy saving that goes from $25.4\%$, if we guarantee thermal comfort, to $49.2\%$, if we allow very little discomfort in terms of rooms' temperature.
\end{enumerate}

%In our previous work~\cite{JainACC2017,Jain2016}, we introduced the concept of DPC for receding horizon control. We now formally present the algorithms for DPC and apply them to a practical building model identified using real weather data and sensor measurements. We present a comprehensive case study to demonstrate the applicability of DPC to demand response, especially how DPC can provide financial incentives to the end-customers bypassing the need for high fidelity models. 
%In particular, we first describe two underlying algorithms: DPC with regression trees and DPC with random forests in Section~\ref{S:dpc}. 
%In Section~\ref{S:proof}, we apply DPC to a bilinear building model, whose parameters were identified using experiments on a building in Switzerland, to show a one-to-one comparison of receding horizon control with DPC against a benchmark MPC controller. In Section~\ref{S:casestudy}, we present a practical application of DPC for Demand Response, where we apply DPC to a multi-story 19 zone building model in EnergyPlus for which model-based control is not economical and practical due to extreme complexity.

@@

%We begin with description of DPC-RT and DPC-En algorithms, and Gaussian Processes for DPC in Section~\ref{S:dpc}. Section~\ref{S:casestudy} presents a case study on DR using a bilinear building model. Here, we also validate the black-box models used for control, and compare DPC against MPC. We discuss the challenges associated with DPC in Section~\ref{SS:challenges}. We conclude the paper with a summary of the results and a brief discussion on the future work in Section~\ref{S:conclusion}.

### 2017-09-29 paper/main.tex
Modified introduction. First complete draft.
edited abstract+intro
Couple of small changes in abstract, intro and keywords
Added flow scheme, motivation. Implemented comments from Italy.

\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{francesco.smarra@univaq.it@univaq.it}

\author[AJaddress,EC]{Achin Jain}
\ead{achinj@seas.unpenn.edu}
\author[TDRaddress,EC]{Tullio de Rubeis}
\ead{tullio.derubeis@graduate.univaq.it}
\author[TDRaddress]{Dario Ambrosini}
\ead{dario.ambrosini@ing.univaq.it}
\author[FSaddress]{Alessandro D'Innocenzo}
\ead{alessandro.dinnocenzo@univaq.it}
\author[AJaddress]{Rahul Mangharam}
\ead{rahulm@seas.upenn.edu}

\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
\address[TDRaddress]{Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
\address[EC]{Equal contribution}

@@

\begin{keyword}
Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real case study
\end{keyword}

>>>>

\author[FSaddress,EC]{Francesco Smarra\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{francesco.smarra@univaq.it}

\author[AJaddress,EC]{Achin Jain}
\ead{achinj@seas.unpenn.edu}
\author[TDRaddress,EC]{Tullio de Rubeis}
\ead{tullio.derubeis@graduate.univaq.it}
\author[TDRaddress]{Dario Ambrosini}
\ead{dario.ambrosini@univaq.it}
\author[FSaddress]{Alessandro D'Innocenzo}
\ead{alessandro.dinnocenzo@univaq.it}
\author[AJaddress]{Rahul Mangharam}
\ead{rahulm@seas.upenn.edu}

\fntext[EC]{Equal contribution}

\address[FSaddress]{Department of Information Engineering, Computer Science and Mathematics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}
\address[AJaddress]{Department of Electrical and Systems Engineering, University of Pennsylvania, Philadelphia, USA}
\address[TDRaddress]{"G. Parolini Lab" - Department of Industrial and Information Engineering and Economics, Universit\`{a} degli Studi dell'Aquila, L'Aquila, Italy}

@@

\begin{keyword}
Energy efficiency, demand response, predictive control, machine learning, EnergyPlus modeling, real application
\end{keyword}
